\pagestyle{plain}
\vspace*{52pt}
\begin{center}
    \large{\textbf{Acknowledgements}}\\[31pt]
\end{center}
First of all, I would like to thank Lincoln University for giving me a chance to prepare the project in partial fulfilment of the requirements of the degree of Master of Computer Science (MCS) awarded by Lincoln University. 
After many months of hard work and sincere effort from my side, this research has been conducted. I would like to acknowledge the following notable personalities who have contributed their valuable efforts in different ways in creation of this research.

\par I would like to give special thanks to Mr. Santosh Dhungana, my research project supervisor for his professional guidance and valuable support and for his useful and constructive recommendations on this project.
I also owe deep gratitude to all reputed authors whose writings have provided me with the necessary guidance and invaluable materials for the enrichment of my research papers in all possible ways. 

\par My special appreciation goes to my colleague and to all my family members, teachers and friends for their continuous encouragement and help to complete this work directly or indirectly.
I wish to thank various people for their contribution to this project: Mr. Bibek Ale Magar and Mr. Dipesh Dulal for their valuable technical support on this project: Mr. Prakash Devkota for their help in collecting data necessary for the project .


% 3 blank lines, 14pt, line spacing 1.5
\vspace{73pt}

\begin{flushleft}
    Signature \\
    Name of the Student: Ujjwol Kayastha\\
    Registration Number: LC0003001469\\
    Date:
\end{flushleft}

\addtocontents{toc}{Acknowledgements\hfill \thepage\par}